---
title: "Problem Set #3"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(broom)
library(modelsummary)
library(estimatr)
library(marginaleffects)
library(readxl)
library(juanr)
```


## Question 1

The North Carolina Rural Health Research Program maintains a list of rural hospital closures in the US since 2005. 

Independent variable: Complete Closure vs Converted Closure 
Dependent Variable: Rural-Urban Commuting Area classification

I hypothesize that the level of rural-ness in regards to hospital location correlates to whether or not the hospital was completely closed or converted. 

```{r}

# Load data 

waste = read_csv("data/2024_us.csv") %>% 
  janitor::clean_names()

new_waste = waste %>% 
  select(x4_facility_name,
         x8_st, 
         x53_5_3_water,
         x65_on_site_release_total, 
         x106_6_2_total_transfer) %>% 
  
   mutate(South = if_else(x8_st %in% c("AL", "AR", "FL", "GA", "KY" ,"LA", "MS",
         "NC", "SC", "TN", "TX", "VA"), 1, 0)) %>%
    filter(!is.na(x53_5_3_water), 
           !is.na(x65_on_site_release_total),
           !is.na(x106_6_2_total_transfer))
new_waste

```


## Question 2 

```{r}

rebel_leader = rebel_leader %>% 
  filter(
    !is.na(dead) ,
    !is.na(death_cause),
    !is.na(birth_year), 
    !is.na(western), 
    !is.na(train_abroad))



```



# Question 3 

```{r}

affairs_clean = affairs |>
  mutate(had_affair = as.numeric(n_affairs > 0)) |>
  select(had_affair, kids, years_married)

```

